---
title: "subjects_descripton_flowchart"
author: "Fernanda Hansen Pacheco de Moraes"
date: "2023-01-16"
output:
  html_document: 
    fig_width: 8
    number_sections: yes
    theme: paper
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: inline
---

# RMARKDOWN AND R SETUP
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
```

```{r working directory}
setwd("~/GitHub/CorticalFolding_AD_Aging")
```

```{r call packages}
library(readr)
library(readxl)
library(tidyverse)
library(lubridate)
library(dtrackr)
library(ggpubr)
```

# set seed for random process

```{r}
set.seed(1)
```

# import subjs info

```{r}
subjects <- read_csv(
  "tabela_sujeitos.csv",
  col_types = cols(
    ...1 = col_skip(),
    age = col_skip(),
    dob = col_skip(),
    SUBJ = col_skip(),
    Session = col_skip(),
    Gender = col_skip(),
    birthday = col_skip(),
    Birthdate = col_skip(),
    NeuroQuant = col_skip(),
    Lipoxina = col_skip()
  )
) %>%
  mutate(SUBJ = str_sub(participant_id, 5, 11),
         Session = as.double(str_sub(session_id, 5, 5)))
```

Translate Diagnostics labels to English

```{r}
subjects$Diagnostic[subjects$Diagnostic == "CONTROLE"] <- "CTL"
subjects$Diagnostic[subjects$Diagnostic == "ALZ"] <- "AD"
subjects$Diagnostic[subjects$Diagnostic == "CCL"] <- "MCI"

subjects$Diagnostic <- factor(subjects$Diagnostic, levels = c("AD", "MCI","CTL"))

```

Merging with morphometrical data to ensure FreeSurfer completition + clinical and CSF data

```{r}
dados_raw <- read_csv("dados_raw.csv") %>%
  filter(hemi == "L", ROI == "hemisphere")  %>% # here I selected results for the left hemisphere and ROI the complete hemisphere
  dplyr::select(c(SUBJ, Session, AvgThickness, ExposedArea))

subjects <- full_join(subjects, dados_raw)
```

Import aseg stats table to include the mean Euler Number (mean between hemispheres).

```{r}
aseg_vol <- read_table("aseg_vol.txt") %>%
  dplyr::select(c(`Measure:volume`, lhSurfaceHoles, rhSurfaceHoles))
```

```{r}
EulerNumber <- read_table("aseg_vol.txt") %>%
  mutate(
    SUBJ = str_split(`Measure:volume`, "/", simplify = TRUE)[, 1],
    LongitudinalCorrection = ifelse(str_sub(
      str_split(`Measure:volume`, "/", simplify = TRUE)[,2], 15, 18
    ) == "long", "yes", "no"),
    Session = as.double(str_sub(str_split(`Measure:volume`, "/", simplify = TRUE)[, 2], 13, 13)),
    lhEulerNumber = 2 - 2 * lhSurfaceHoles,
    rhEulerNumber = 2 - 2 * rhSurfaceHoles,
    meanEulerNumber = (lhEulerNumber + rhEulerNumber) / 2
  ) %>%
  filter(LongitudinalCorrection == "yes") %>%
  dplyr::select(c(SUBJ, Session, lhEulerNumber, rhEulerNumber, meanEulerNumber)) %>%
  unique()
  
subjects <- full_join(subjects, EulerNumber) %>%
  unique()

medianEulerNumber <- subjects %>%
  group_by(Diagnostic) %>%
  mutate(median_EulerNumber = median(meanEulerNumber)) %>%
  unique()

subjects <- full_join(subjects, medianEulerNumber) %>%
  mutate(meanEulerNumber_norm = sqrt(-1*(meanEulerNumber - median_EulerNumber))) %>%
  unique()

```

# Euler number

```{r}
ggplot(subjects, aes(x = SUBJ, y = meanEulerNumber, color = Diagnostic)) + 
  geom_point() +
  theme_pubr()

ggplot(subjects, aes(x = Diagnostic, y = meanEulerNumber, color = Diagnostic)) + 
  geom_boxplot() +
  theme_pubr()
```

# Euler number before surface corretion

```{r}
eulernumber_befrsurfacecorrection <-
  read_csv("eulernumber_ad6.csv") %>%
  mutate(
    SUBJ = str_split(fullpath, pattern = "/", simplify = TRUE)[, 8],
    Session = as.double(str_split(SUBJ, pattern = "_ses-", simplify = TRUE)[, 2]),
    SUBJ = str_split(SUBJ, pattern = "_ses-", simplify = TRUE)[, 1]) %>%
  filter(!is.nan(lhEuler),
         SUBJ != str_detect(SUBJ, pattern = "_tmpl")) %>%
  dplyr::select(-c(fullpath))

eulernumber_befrsurfacecorrection <- eulernumber_befrsurfacecorrection %>%
  mutate(
    meanEulernumber_befrsurfacecorrection = (lhEuler+rhEuler)/2
  )

# eulernumber_renumbered_lh <- eulernumber_renumbered %>%
#   dplyr::select(SUBJ, lhEuler, meanEulernumber_befrsurfacecorrection) %>%
#   mutate(hemi = "left")
# colnames(eulernumber_renumbered_lh)[2] <- "Eulernumber"
# eulernumber_renumbered_rh <- eulernumber_renumbered %>%
#   dplyr::select(SUBJ, rhEuler, meanEulernumber_befrsurfacecorrection) %>%
#   mutate(hemi = "right")
# colnames(eulernumber_renumbered_rh)[2] <- "Eulernumber"
# eulernumber_befrsurfacecorrection <- rbind(eulernumber_renumbered_lh, eulernumber_renumbered_rh)

subjects_euler <- full_join(subjects, eulernumber_befrsurfacecorrection) %>%
  filter(!is.na(Diagnostic), Session == 1)
```

```{r}
ggplot(subjects_euler, aes(x = Diagnostic, y = meanEulernumber_befrsurfacecorrection)) +
  geom_boxplot(outlier.color = "red") +
  theme_pubr()

subjects_euler %>%
  summarise(
    mean = mean(meanEulernumber_befrsurfacecorrection),
    sd = sd(meanEulernumber_befrsurfacecorrection)
  )

subjects_euler %>%
  group_by(Diagnostic) %>%
  summarise(
    mean = mean(meanEulernumber_befrsurfacecorrection),
    sd = sd(meanEulernumber_befrsurfacecorrection)
  )

summary(aov(meanEulernumber_befrsurfacecorrection ~ Diagnostic, data = subjects_euler))
TukeyHSD(aov(meanEulernumber_befrsurfacecorrection ~ Diagnostic, data = subjects_euler))

subjects_euler %>%
  filter(meanEulernumber_befrsurfacecorrection < -100) %>%
  dplyr::select(SUBJ, Age, Diagnostic, meanEulernumber_befrsurfacecorrection) %>%
  unique()
```

Centering

```{r}
mean <- subjects_euler %>%
  filter(Diagnostic == "CTL") %>%
  summarise(
    mean = mean(meanEulernumber_befrsurfacecorrection),
    sd = sd(meanEulernumber_befrsurfacecorrection)
  )

subjects_euler <- cbind(subjects_euler, mean) %>%
  mutate(z_euler = (meanEulernumber_befrsurfacecorrection- mean)/sd,
         rutherford_euler = sqrt(z_euler*(-1)))

ggplot(subjects_euler, aes(x = Diagnostic, y = rutherford_euler)) +
  geom_boxplot() +
  geom_hline(yintercept = 10) +
  theme_pubr()
```

# Visual quality Control

```{r}
VisualQC_ses_1  <- read_excel("Verificacao qualidade segmentacao FS Zika e CCD 2.xlsx",
    sheet = "AD - IDOR ses 1 long", col_types = c("text", 
        "text", "text", "skip", "numeric", 
        "skip", "skip", "skip", "skip", "skip", 
        "skip", "skip")) %>%
  mutate(Session = as.double(str_sub(Session,5,5)),
          Classification = Fernanda) %>%
  dplyr::select(-c(Fernanda))

VisualQC_ses_2 <- read_excel("Verificacao qualidade segmentacao FS Zika e CCD 2.xlsx", sheet = "AD - IDOR ses 2 long", col_types = c("text", 
        "text", "skip", "numeric", "skip", 
        "skip", "skip", "skip", "skip")) %>%
   mutate(SUBJ = str_sub(SUBJ,5,11),
          Session = as.double(str_sub(Session,5,5)),
          Classification = Fernanda,
          `Longitudinal correction` = "long") %>%
  dplyr::select(-c(Fernanda))

VisualQC_ses_3 <- read_excel("Verificacao qualidade segmentacao FS Zika e CCD 2.xlsx", sheet = "AD - IDOR ses 3 long", col_types = c("text", 
        "text", "skip", "numeric", "skip", 
        "skip", "skip", "skip", "skip")) %>%
   mutate(SUBJ = str_sub(SUBJ,1,8),
          Session = as.double(str_sub(Session,5,5)),
          Classification = Fernanda,
          `Longitudinal correction` = "long") %>%
  dplyr::select(-c(Fernanda))


VisualQC <- full_join(VisualQC_ses_1, VisualQC_ses_2)
# %>%
#   full_join(VisualQC_ses_3)
         
subjects <- full_join(subjects, VisualQC)

```

## Prioridades Ananssa

```{r}
prioridades_simoa <- read_excel("C:/Users/ferna/Downloads/Amostras usadas SIMOA_Prioridades para neuroimagem_2.xlsx", 
    col_types = c("skip", "text", "text", 
        "date", "numeric", "skip", "skip")) %>%
  mutate(acq_date = Acquisition_Date) %>%
  select(-c(Acquisition_Date, Diag.))
```


```{r}

infos <- full_join(subjects, prioridades_simoa) %>%
  select(c(SUBJ, Session, acq_date, AvgThickness, Classification, Prioridade))
write.csv(infos, "prioridades_simoa.csv")

```


```{r}
write.csv(subjects, "subjects.csv")

subjects %>%
  filter(Session == 1, Classification < 1 ) %>%
  dplyr::select(SUBJ, Age, Diagnostic, Classification) %>%
  unique()
```


# inclusion/exclusion criteria

All subjects included must have:
- been scanned at Philips Achieva 3T (to unify the acquisition setup);
- Diagnostic equal to Healthy Controls, Mild Cognitive Impairment and Alzheimer's Disease, without any other diagnostic;
- As a cross-sectional study, we selected the first acquisition (ses-1), if there are more than one;
- Subjects with processing errors on FreeSurfer (meaning here no results ExposedArea) were excluded.

```{r}
dataset1 = subjects %>%
  track("Starts with {.count} acquistions") %>%
  include_any(machine == "Philips-Achieva"  ~ "{.included} images acquired with Philips Achieva 3T") %>%
  include_any(Session == "1"  ~ "{.included} first sessions only") %>%
  include_any(
    Diagnostic == "CTL" |
      Diagnostic == "MCI" |
      Diagnostic == "AD"  ~ "{.included} subjects with the selected diagnostics (CTL, MCI or AD)"
  ) %>%
  group_by(Diagnostic) %>%
  comment("{Diagnostic} has {.count} subjects") %>%
  include_any(ESC == 8 |
                ESC > 8  ~ "{.included} years of education => 8") %>%
  include_any(!is.na(ExposedArea) ~ "{.included} processed with FreeSurfer") %>%
  exclude_all(Classification < 1  ~ "{.excluded} rejected on Visual QC") %>%
  ungroup(.messages = "Ends with {.total} subjects")

dataset1 %>% flowchart()

write.csv(dataset1, "dataset1.csv")
```

```{r}
# subj_names <- dataset1 %>% dplyr::select(c(SUBJ)) 
# subj_names <- lapply(subj_names,sort)
# subj_names
```

## Quality control issues

```{r}

dataset1 %>%
  group_by(Diagnostic,
           Classification) %>%
  summarise(
    N = n_distinct(SUBJ),
    age = paste(signif(mean(Age), 2), "±", signif(sd(Age), 2)),
    age_range = paste(signif(min(Age), 2), "; ", signif(max(Age), 2))
  )

dataset1 %>%
  group_by(`brainmask error`,
           `underestimation temporal region`) %>%
  summarise(
    N_percent = round(n_distinct(SUBJ)*100/121,1),
    age = paste(signif(mean(Age), 2), "±", signif(sd(Age), 2)),
    age_range = paste(signif(min(Age), 2), "; ", signif(max(Age), 2))
  )

```

