---
title: "AD_ADNI_IDOR"
author: "Fernanda Hansen P. de Moraes"
date: "23/06/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
```

```{r preparo, message=FALSE, warning=FALSE, include=FALSE}
# PREPARO

# define a area de trabalho
setwd("C:/Users/ferna/Documents/idor/Gyrification/RRRRRR/comparing_samples/") # se no computador

# carrega os pacotes utilizados
source("01-call_packages.R")

# Chama funcoes caso necessário
source("02-funcoes.R")

# onde estao os arquivos de resultados:
path <- str_c("C:/Users/ferna/Documents/idor/Gyrification/STATS/long_tables/")

path_yujiangscript <- str_c("C:/Users/ferna/Documents/idor/Gyrification/STATS/Wang_code_extraction/")

path_yujiangscript_newsurfaces <- str_c("C:/Users/ferna/Documents/idor/Gyrification/STATS/Wang_code_extraction/")
  
path_lobes <- str_c("C:/Users/ferna/Documents/idor/Gyrification/STATS/LobesScaling/")
 
path_sessions <- str_c("C:/Users/ferna/Documents/idor/Gyrification/Processamento_longitudinal/V6/data/")

Age.cor = 25

# Prepara os dados para analise
source("05-analises_prep_ADNI.R")

dados_datasetscomp <- dados_datasetscomp %>%
   filter(Sample == "ADNI" | Sample == "IDOR")  %>% 
   # dplyr::select(c(machine, SUBJ, Age, Gender, Diagnostic, Sample, ROI, hemi, AvgThickness, TotalArea, ExposedArea, GMvolume, logAvgThickness, logTotalArea, logExposedArea, localGI, k, K, S, I, Knorm, Snorm, Inorm, c, TotalArea_corrected, ExposedArea_corrected, logTotalArea_corrected, logExposedArea_corrected, localGI_corrected, K_corrected, S_corrected, I_corrected, APOE4, MMSE, FLDSTRENG, FSVERSION, ABETA, TAU, PTAU, ABETA, localGI_age_decay, localGI_shiftc, localGI_age_decay_shiftc, K_age_decay, K_shiftc, K_age_decay_shiftc, I_age_decay, I_shiftc, I_age_decay_shiftc, S_age_decay, S_shiftc, S_age_decay_shiftc, Knorm_shiftc, Snorm_shiftc, Inorm_shiftc, Knorm_age_decay, Snorm_age_decay, Inorm_age_decay, Knorm_age_decay_shiftc, Snorm_age_decay_shiftc, Inorm_age_decay_shiftc, Age_interval10, Age_interval)) %>%
   mutate(
     ABETA = ifelse(Sample == "ADNI", ABETA, ABETA),
     TAU = as.double(TAU))

dados_datasetscomp$Diagnostic <- factor(dados_datasetscomp$Diagnostic, levels = c("CTL", "MCI", "AD"))

dados_ADNI <- dados_datasetscomp %>%
  filter(Sample == "ADNI", ROI == "hemisphere") %>%
  dplyr::select(c(SUBJ, Age, Gender, Diagnostic, Sample, ROI, hemi, AvgThickness, TotalArea, ExposedArea, GMvolume, k, K, S, I, MMSE, ABETA, TAU, Age_interval10, Age_interval)) 

# write.csv(dados_ADNI, "dados_ADNI.csv")
```

```{r functions, message=FALSE, warning=FALSE}
## define functions

# test angular coeficinet versus theoretical value
test_coef <- function(reg, coefnum, val){
  co <- coef(summary(reg))
  tstat <- (co[coefnum,1] - val)/co[coefnum,2]
  2 * pt(abs(tstat), reg$df.residual, lower.tail = FALSE)
}

# wrap text
wrapper <- function(x, ...) paste(strwrap(x, ...), collapse = "\n")
```

```{r} 
# COLOR BLIND PALETTE WITH BLACK
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbbPalette2 <- c("#D55E00", "#E69F00", "#56B4E9", "#0072B2", "#CC79A7", "#009E73", "#F0E442")
```

# set seed for random process

```{r}
set.seed(1)
```

# Data description

```{r}
filter(dados_datasetscomp, ROI == "hemisphere") %>%
  group_by(Diagnostic, Sample) %>%
  summarise(
    N = n_distinct(SUBJ),
    age = paste(signif(mean(Age), 2), "±", signif(sd(Age), 2)),
#   age_range = paste(signif(min(Age), 2), "; ", signif(max(Age), 2)),
    AvgT = paste(signif(mean(AvgThickness), 2), "±", signif(sd(AvgThickness), 2)),
AT = paste(signif(mean(TotalArea), 2), "±", signif(sd(TotalArea), 2)),
AE = paste(signif(mean(ExposedArea), 2), "±", signif(sd(ExposedArea), 2))
  )

filter(dados_datasetscomp, ROI == "hemisphere") %>%
  group_by(Diagnostic, Sample) %>%
  summarise(
    N = n_distinct(SUBJ),
#   age_range = paste(signif(min(Age), 2), "; ", signif(max(Age), 2)),
    k = paste(signif(mean(k), 2), "±", signif(sd(k), 2)),
    K = paste(signif(mean(K), 2), "±", signif(sd(K), 2)),
    S = paste(signif(mean(S), 2), "±", signif(sd(S), 2)),
    I = paste(signif(mean(I), 2), "±", signif(sd(I), 2)),
  )

filter(dados_datasetscomp, ROI == "hemisphere") %>%
  group_by(Diagnostic, Sample, Gender) %>%
  summarise(
    N = n_distinct(SUBJ)
  ) %>% kable(digits = 2) %>% kable_styling()

filter(dados_datasetscomp, ROI == "hemisphere") %>%
  filter(!is.na(TAU), !is.na(ABETA)) %>%
  group_by(Diagnostic, Sample) %>%
  summarise(
    N = n_distinct(SUBJ),
    age = paste(signif(mean(Age), 2), "±", signif(sd(Age), 2)),
    TAU = paste(signif(mean(TAU, na.rm = TRUE), 2), "±", signif(sd(TAU, na.rm = TRUE), 2)),
    ABETA = paste(signif(mean(ABETA, na.rm = TRUE), 2), "±", signif(sd(ABETA, na.rm = TRUE), 2))
  )
```
```{r}
dados_v1 <- filter(dados_datasetscomp, ROI == "F" | ROI == "T" | ROI == "O" | ROI == "P" | ROI == "hemisphere") %>%
  droplevels()

dados <- filter(dados_datasetscomp, ROI == "F" | ROI == "T" | ROI == "O" | ROI == "P" | ROI == "hemisphere", Sample == "ADNI") %>%
  droplevels()

# lobe data
dados_lobos_v1 <- unique(filter(dados_datasetscomp, ROI == "F" | ROI == "T" | ROI == "O" | ROI == "P",  !is.na(K_age_decay), SUBJ != "SUBJ211", SUBJ != "SUBJ223")) %>%
  droplevels()

dados_lobos_v1_ADNI <- unique(filter(dados_datasetscomp, ROI == "F" | ROI == "T" | ROI == "O" | ROI == "P",  !is.na(K_age_decay), SUBJ != "SUBJ211", SUBJ != "SUBJ223", Sample == "ADNI")) %>%
  droplevels()

# hemisphere data
dados_hemi_v1 <- unique(filter(dados_datasetscomp, ROI == "hemisphere", !is.na(K_age_decay)))

dados_hemi_v1_ADNI <- unique(filter(dados_datasetscomp, ROI == "hemisphere", Sample == "ADNI", !is.na(K_age_decay)))
```

### Stats tests 

#### Age
```{r}
summary(aov(Age ~ Diagnostic, data = dados_hemi_v1_ADNI))
TukeyHSD(aov(Age ~ Diagnostic, data = dados_hemi_v1_ADNI))
```

#### T

```{r}
summary(aov(AvgThickness ~ Diagnostic, data = dados_hemi_v1_ADNI))
TukeyHSD(aov(AvgThickness ~ Diagnostic, data = dados_hemi_v1_ADNI))
```

#### AT

```{r}
summary(aov(TotalArea ~ Diagnostic, data = dados_hemi_v1_ADNI))
TukeyHSD(aov(TotalArea ~ Diagnostic, data = dados_hemi_v1_ADNI))
```

#### AE

```{r}
summary(aov(ExposedArea ~ Diagnostic, data = dados_hemi_v1_ADNI))
TukeyHSD(aov(ExposedArea ~ Diagnostic, data = dados_hemi_v1_ADNI))
```

#### k

```{r}
summary(aov(k ~ Diagnostic, data = dados_hemi_v1_ADNI))
TukeyHSD(aov(k ~ Diagnostic, data = dados_hemi_v1_ADNI))
```

#### K

```{r}
summary(aov(K ~ Diagnostic, data = dados_hemi_v1_ADNI))
TukeyHSD(aov(K ~ Diagnostic, data = dados_hemi_v1_ADNI))
```

#### S

```{r}
summary(aov(S ~ Diagnostic, data = dados_hemi_v1_ADNI))
TukeyHSD(aov(S ~ Diagnostic, data = dados_hemi_v1_ADNI))
```

#### I

```{r}
summary(aov(I ~ Diagnostic, data = dados_hemi_v1_ADNI))
TukeyHSD(aov(I ~ Diagnostic, data = dados_hemi_v1_ADNI))
```

## Cortical Folding Model - ADNI

```{r}
summary(lm(
  1 / 2 * log10(AvgThickness) + log10(TotalArea) ~ log10(ExposedArea),
  data = dados_hemi_v1_ADNI,
  na.action = na.omit
))

# Displays confidence interval
tidy(lm(
  1 / 2 * log10(AvgThickness) + log10(TotalArea) ~ log10(ExposedArea),
  data = dados_hemi_v1_ADNI,
  na.action = na.omit
), conf.int = TRUE)

paste(
  "Student's t test comapring slope with theoretical value 1.25. t = ",
  signif(abs(coef(summary(
    lm(
      1 / 2 * log10(AvgThickness) + log10(TotalArea) ~ log10(ExposedArea),
      data = dados_hemi_v1_ADNI,
      na.action = na.omit
    )
  ))[2, 1] - 1.25) / coef(summary(
    lm(
      1 / 2 * log10(AvgThickness) + log10(TotalArea) ~ log10(ExposedArea),
      data = dados_hemi_v1_ADNI,
      na.action = na.omit
    )
  ))[2, 2], 3) 
)
paste(
  "Student's t test comapring slope with theoretical value 1.25. p value = ",
  signif(test_coef(
    lm(
      1 / 2 * log10(AvgThickness) + log10(TotalArea) ~ log10(ExposedArea),
      data = dados_hemi_v1_ADNI,
      na.action = na.omit
    ),
    2,
    1.25
  ),3)
)
```

```{r}
dados_hemi_v1_ADNI %>%
  group_by(Diagnostic) %>%
  do(fit_Age = tidy(
    lm(
      1 / 2 * log10(AvgThickness) +  log10(TotalArea) ~  log10(ExposedArea),
      data = .,
      na.action = na.omit
    ),
    conf.int = TRUE
  )) %>%
  unnest(cols = c(fit_Age))

dados_hemi_v1_ADNI %>%
  group_by(Diagnostic) %>%
  do(fit_Age = glance(
    lm(
      1 / 2 * log10(AvgThickness) +  log10(TotalArea) ~  log10(ExposedArea),
      data = .,
      na.action = na.omit
    ),
    conf.int = TRUE
  )) %>%
  unnest(cols = c(fit_Age))
```


## Correlation within cortical fodling variables and age healthy subjects)

### Slope alpha

```{r}
lm_Age <-
  filter(
    dados_hemi_v1_ADNI,
    Diagnostic == "CTL") %>%
  group_by(Age_interval) %>%
  do(fit_Age = tidy(
    lm(
      1 / 2 * log10(AvgThickness) +  log10(TotalArea) ~  log10(ExposedArea),
      data = .,
      na.action = na.omit
    ),
    conf.int = TRUE
  )) %>%
  unnest(cols = c(fit_Age))

N_subj <-
  filter(
    dados_hemi_v1_ADNI,
    Diagnostic == "CTL"
    ) %>%
  group_by(Age_interval) %>%
  summarise(N_SUBJ = n_distinct(SUBJ))

N_subj

fig1_adni <- ggplot(data = filter(lm_Age, term == "log10(ExposedArea)", Age_interval != "[90,95)", Age_interval != "[95,100]"),
                aes(x = Age_interval,
                    y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error)) +
  geom_hline(yintercept = 1.25, linetype = "dashed") +
  geom_text(aes(label = filter(N_subj, Age_interval != "[90,95)", Age_interval != "[95,100]")$N_SUBJ), nudge_y = 0.3) +
  theme_pubr() +
  labs(y = "Slope", x = "Age [years]") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

fig1_adni
ggsave("fig1_adni.pdf", plot = fig1_adni, width = 9, height = 11.37, units = "cm", device = "pdf")

lm_Age <- lm_Age %>% mutate(Age_interval = as.double((str_sub(Age_interval,2,3))))

cor.test(filter(lm_Age, term == "log10(ExposedArea)", Age_interval != "[90,95)", Age_interval != "[95,100]" )$estimate, filter(lm_Age, term == "log10(ExposedArea)", Age_interval != "[90,95)", Age_interval != "[95,100]")$Age_interval)
```

### K

```{r}
 ggplot(data = dados_hemi_v1_ADNI, aes(Age, K, color = Diagnostic, fill = Diagnostic, alpha = 0.4)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  theme_pubr() +
    guides(alpha = "none", color = "none", fill = "none") + 
  labs(x = "Age [years]") +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)

cor.test(filter(dados_hemi_v1_ADNI, Diagnostic == "CTL")$K, filter(dados_hemi_v1_ADNI, Diagnostic == "CTL")$Age)

dados_hemi_v1_ADNI %>%
  group_by(Diagnostic) %>%
  do(fit_Age = tidy(
    lm(
      K ~  Age,
      data = .,
      na.action = na.omit
    ),
    conf.int = TRUE
  )) %>%
  unnest(cols = c(fit_Age))

dados_hemi_v1_ADNI %>%
  group_by(Diagnostic) %>%
  do(fit_Age = glance(
    lm(
      K ~  Age,
      data = .,
      na.action = na.omit
    ),
    conf.int = TRUE
  )) %>%
  unnest(cols = c(fit_Age))

```


### S

```{r}

 ggplot(data = dados_hemi_v1_ADNI, aes(Age, S, color = Diagnostic, fill = Diagnostic, alpha = 0.4)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  theme_pubr() +
    guides(alpha = "none", color = "none", fill = "none") + 
  labs(x = "Age [years]") +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)

cor.test(filter(dados_hemi_v1_ADNI, Diagnostic == "CTL")$S, filter(dados_hemi_v1_ADNI, Diagnostic == "CTL")$Age)
```
### I

```{r}

 ggplot(data = dados_hemi_v1_ADNI, aes(Age, I, color = Diagnostic, fill = Diagnostic, alpha = 0.4)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  theme_pubr() +
    guides(alpha = "none", color = "none", fill = "none") + 
  labs(x = "Age [years]") +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)

cor.test(filter(dados_hemi_v1_ADNI, Diagnostic == "CTL")$I, filter(dados_hemi_v1_ADNI, Diagnostic == "CTL")$Age)
```

**FIGURE 1**
```{r}
fig2a <- ggplot(dados_hemi_v1_ADNI,
                aes(
                  log10(ExposedArea),
                  log10(sqrt(AvgThickness) * TotalArea),
                  color = Diagnostic,
                  fill = Diagnostic,
                  alpha = 0.4
                )) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  # geom_abline(slope = 1.25,
  #             intercept = coef(lm(
  #               log10(sqrt(AvgThickness) * TotalArea) ~ log10(ExposedArea),
  #               data = dados_hemi_v1_ADNI,
  #               na.action = na.omit
  #             ))[1],
  #             color = "black") +
  labs(x = expression('log'[10] * 'A'['E']),
       y = expression('log'[10] * 'A'['T'] * sqrt('T'))) +
  guides(alpha = "none") +
  theme_pubr() +
  scale_x_continuous(limits = c(4.45, 4.65)) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)
```

```{r}
fig1b <- ggplot(data = dados_hemi_v1_ADNI, aes(Age, K, color = Diagnostic, fill = Diagnostic, alpha = 0.4)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  theme_pubr() +
    guides(alpha = "none", color = "none", fill = "none") + 
  labs(x = "Age [years]") +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)
```

```{r}
fig1_alt_2 <-
  ggarrange(
    fig2a,
    fig1b,
    labels = c("A", "B"),
    ncol = 1,
    common.legend = TRUE,
    legend = "bottom"
  )

ggsave("fig1_alt_2_adni.pdf", plot = fig1_alt_2, width = 9, height = 11.37, units = "cm", device = "pdf")
# ggsave("fig1_alt_2.png", plot = fig1_alt_2, width = 9, height = 11.37, units = "cm", device = "png")
```

```{r figure1, fig.height=3.54, fig.width=4.48}

fig1_alt_2

```

## Diagnostic discrimination and prediction.

### K difference

```{r}
t.test(K ~ Diagnostic, data = dados_hemi_v1_ADNI)
```

### K is reduced with age, as cortical thickness, total area and exposed area

K decrease with age is shown on Figure 1 B. Cortical Thickness, Total area and Exposed area:

```{r}
T <- ggplot(data = dados_hemi_v1_ADNI, aes(Age, AvgThickness, color = Diagnostic, fill = Diagnostic, alpha = 0.4)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  theme_pubr() +
    guides(alpha = "none", color = "none", fill = "none") + 
  labs(x = "Age [years]") +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)

AT <- ggplot(data = dados_hemi_v1_ADNI, aes(Age, TotalArea, color = Diagnostic, fill = Diagnostic, alpha = 0.4)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  theme_pubr() +
    guides(alpha = "none", color = "none", fill = "none") + 
  labs(x = "Age [years]", y = "Total Area 10^-5 ") +
  scale_y_continuous(
    labels = function(x)
      x / 10000) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)

AE <- ggplot(data = dados_hemi_v1_ADNI, aes(Age, ExposedArea, color = Diagnostic, fill = Diagnostic, alpha = 0.4)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  theme_pubr() +
    guides(alpha = "none", color = "none", fill = "none") + 
  labs(x = "Age [years]", y = "Exposed Area 10^-5 ") +
  scale_y_continuous(
    labels = function(x)
      x / 10000) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)

```

```{r}
ggarrange(T, AT, AE, ncol = 1, common.legend = TRUE, legend = "bottom")
```

### Optimal cut-offs
#### K
```{r echo=FALSE}
cpK <-
    cutpointr(
        filter(dados_hemi_v1_ADNI, Diagnostic == "AD" | Diagnostic == "CTL"),
        K,
        Diagnostic,
        pos_class = "AD",
        neg_class = "CTL",
        method = maximize_boot_metric,
        metric = sum_sens_spec,
        na.rm = TRUE,
        boot_runs = 1000,
    use_midpoints = TRUE)
summary(cpK)
plot(cpK)

```

##### Sex as subgroups

```{r echo=FALSE}
cpK_sex <-
    cutpointr(
        filter(dados_hemi_v1_ADNI, Diagnostic == "AD" | Diagnostic == "CTL"),
        K,
        Diagnostic,
        Gender,
        pos_class = "AD",
        neg_class = "CTL",
        method = maximize_boot_metric,
        metric = sum_sens_spec,
        na.rm = TRUE,
        boot_runs = 1000,
    use_midpoints = TRUE)
summary(cpK_sex)

```

#### log Avg Thickness
```{r}
cpT <-
    cutpointr(
        filter(dados_hemi_v1_ADNI, Diagnostic == "AD" | Diagnostic == "CTL"),
       logAvgThickness,
        Diagnostic,
        pos_class = "AD",
        neg_class = "CTL",
        method = maximize_boot_metric,
        metric = sum_sens_spec,
        na.rm = TRUE,
        boot_runs = 1000,
    use_midpoints = TRUE)
summary(cpT)

```

##### Sex as subgroups

```{r echo=FALSE}
cpT_sex <-
    cutpointr(
        filter(dados_hemi_v1, Diagnostic == "AD" | Diagnostic == "CTL"),
       logAvgThickness,
        Diagnostic,
        Gender,
        pos_class = "AD",
        neg_class = "CTL",
        method = maximize_boot_metric,
        metric = sum_sens_spec,
        na.rm = TRUE,
        boot_runs = 1000,
    use_midpoints = TRUE)
summary(cpT_sex)

```

```{r}
lab1 = paste("AD: ACC=", signif(cpK$acc,2),"\nSENS=",signif(cpK$sensitivity,2),"\nSPEC=",signif(cpK$specificity,2)
             # ,"\nMCI: ACC=", signif(cpK_MCI$acc,2),"\nSENS=",signif(cpK_MCI$sensitivity,2),"\nSPEC=",signif(cpK_MCI$specificity,2)
             )

xrng1 <- range(dados_hemi_v1_ADNI$K)

cutpoint_a <- ggplot(dados_hemi_v1_ADNI, aes(x = K, color = Diagnostic, fill = Diagnostic, alpha = 0.4))+
    geom_density() +
    geom_vline(xintercept = cpK$optimal_cutpoint, linetype = "dashed") + 
    # geom_vline(xintercept = cpK_MCI$optimal_cutpoint, linetype = "dotted") + 
    theme_pubr() +
    guides(alpha = "none") +
    theme(axis.title = element_text(size = 11),
          axis.text = element_text(size = 10), text = element_text(size = 10)) +
    scale_x_continuous(
        labels = scales::number_format(accuracy = 0.01), limits=c(-0.64,-0.48)
        # , breaks = c(-0.62, -0.60, -0.58, -0.56, -0.52, -0.49)
        ) +
    scale_fill_manual(values=cbbPalette2) +
    scale_colour_manual(values=cbbPalette2)
# +
#     annotate("text", x = xrng1[1], y = Inf, vjust = 1.1, label = lab1, size = 2)

lab2 = paste("AD: ACC=", signif(cpT$acc,2),"\nSENS=",signif(cpT$sensitivity,2),"\nSPEC=",signif(cpT$specificity,2)
             # ,"\nMCI: ACC=", signif(cpT_MCI$acc,2),"\nSENS=",signif(cpT_MCI$sensitivity,2),"\nSPEC=",signif(cpT_MCI$specificity,2)
             )

xrng2 <- range(dados_hemi_v1_ADNI$logAvgThickness)

cutpoint_b <- ggplot(dados_hemi_v1_ADNI, aes(x = logAvgThickness, color = Diagnostic, fill = Diagnostic, alpha = 0.4))+
    geom_density() +
    geom_vline(xintercept = cpT$optimal_cutpoint, linetype = "dashed") + 
    # geom_vline(xintercept = cpT_MCI$optimal_cutpoint, linetype = "dotted") + 
    theme_pubr() +
    guides(alpha = "none") +
    theme(axis.title = element_text(size = 11),
          axis.text = element_text(size = 10), text = element_text(size = 10), legend.position = "none") +
    labs(x = expression('log'[10]*'T')) +
    labs(x = expression('log'[10]*'T')) + scale_x_continuous(limits=c(0.24,0.48), n.breaks = 6) +
    scale_fill_manual(values=cbbPalette2) +
    scale_colour_manual(values=cbbPalette2)
# +
#     annotate("text", x = xrng2[1], y = Inf, vjust = 1.1, label = lab2, size = 2)
# cutpoint_b

fig_cutpoint <- ggarrange(cutpoint_a, cutpoint_b, labels = c("A", "B"),  ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")
```

```{r}
fig_cutpoint
```

#### GI
```{r}
cpGI <-
    cutpointr(
        filter(dados_hemi_v1, Diagnostic == "AD" | Diagnostic == "CTL"),
       localGI,
        Diagnostic,
        pos_class = "AD",
        neg_class = "CTL",
        method = maximize_boot_metric,
        metric = sum_sens_spec,
        na.rm = TRUE,
        boot_runs = 1000,
    use_midpoints = TRUE)
summary(cpGI)

```

##### Sex as subgroups

```{r}
cpGI_sex <-
    cutpointr(
        filter(dados_hemi_v1, Diagnostic == "AD" | Diagnostic == "CTL"),
       localGI,
        Diagnostic,
       Gender,
        pos_class = "AD",
        neg_class = "CTL",
        method = maximize_boot_metric,
        metric = sum_sens_spec,
        na.rm = TRUE,
        boot_runs = 1000,
    use_midpoints = TRUE)
summary(cpGI_sex)

cpGI_MCI_sex <-
    cutpointr(
        filter(dados_hemi_v1, Diagnostic == "MCI" | Diagnostic == "CTL"),
       localGI,
        Diagnostic,
       Gender,
        pos_class = "MCI",
        neg_class = "CTL",
        method = maximize_boot_metric,
        metric = sum_sens_spec,
        na.rm = TRUE,
        boot_runs = 1000,
    use_midpoints = TRUE)
summary(cpGI_MCI_sex)
```


#### Deaged
##### K
```{r}
cpK_deaged <-
    cutpointr(
        filter(dados_hemi_v1_ADNI, Diagnostic == "AD" | Diagnostic == "CTL"),
        K_age_decay,
        Diagnostic,
        pos_class = "AD",
        neg_class = "CTL",
        method = maximize_boot_metric,
        metric = sum_sens_spec,
        na.rm = TRUE,
        boot_runs = 1000,
    use_midpoints = TRUE)
summary(cpK_deaged)
plot(cpK_deaged)

```

##### log Avg Thickness

```{r}
cpT_deaged <-
    cutpointr(
        filter(dados_hemi_v1_ADNI, Diagnostic == "AD" | Diagnostic == "CTL"),
       logAvgThickness_age_decay,
        Diagnostic,
        pos_class = "AD",
        neg_class = "CTL",
        method = maximize_boot_metric,
        metric = sum_sens_spec,
        na.rm = TRUE,
        boot_runs = 1000,
    use_midpoints = TRUE)
summary(cpT_deaged)

```

##### GI
```{r}
cpGI_deaged <-
    cutpointr(
        filter(dados_hemi_v1, Diagnostic == "AD" | Diagnostic == "CTL"),
        localGI_age_decay,
        Diagnostic,
        pos_class = "AD",
        neg_class = "CTL",
        method = maximize_boot_metric,
        metric = sum_sens_spec,
        na.rm = TRUE,
        boot_runs = 1000,
    use_midpoints = TRUE)
summary(cpGI_deaged)
# plot(cpK)

```

```{r}
lab3 = paste("AD: ACC=", signif(cpK_deaged$acc,2),"\nSENS=",signif(cpK_deaged$sensitivity,2),"\nSPEC=",signif(cpK_deaged$specificity,2)
             # ,"\nMCI: ACC=", signif(cpK_MCI_deaged$acc,2),"\nSENS=",signif(cpK_MCI_deaged$sensitivity,2),"\nSPEC=",signif(cpK_MCI_deaged$specificity,2)
             )

cutpoint_a_deaged <- ggplot(dados_hemi_v1_ADNI, aes(x = K_age_decay, color = Diagnostic, fill = Diagnostic, alpha = 0.4))+
  geom_density() +
  geom_vline(data = cpK_deaged, aes(xintercept = optimal_cutpoint), linetype = "dashed") + 
    # geom_vline(data = cpK_MCI_deaged, aes(xintercept = optimal_cutpoint), linetype = "dotted") + 
  theme_pubr() +
  guides(alpha = "none", linetype = "none") +
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10), text = element_text(size = 10)) +
  labs(x = "K (After age correction)")  + scale_x_continuous(
              labels = scales::number_format(accuracy = 0.01), limits=c(-0.64,-0.48)
              # , breaks = c(-0.62, -0.6, -0.58, -0.55, -0.52, -0.49)
              ) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2) 
# +
#     annotate("text", x = xrng1[1], y = Inf, vjust = 1.1, label = lab3, size = 2)
# cutpoint_a_deaged

lab4 = paste("AD: ACC=", signif(cpT_deaged$acc,2),"\nSENS=",signif(cpT_deaged$sensitivity,2),"\nSPEC=",signif(cpT_deaged$specificity,2)
             # ,"\nMCI: ACC=", signif(cpT_MCI_deaged$acc,2),"\nSENS=",signif(cpT_MCI_deaged$sensitivity,2),"\nSPEC=",signif(cpT_MCI_deaged$specificity,2)
             )

cutpoint_b_deaged <- ggplot(dados_hemi_v1_ADNI, aes(x = logAvgThickness_age_decay, color = Diagnostic, fill = Diagnostic, alpha = 0.4))+
  geom_density() +
  geom_vline(data = cpT_deaged, aes(xintercept = optimal_cutpoint), linetype = "dashed") + 
  # geom_vline(data = cpT_MCI_deaged, aes(xintercept = optimal_cutpoint), linetype = "dotted") + 
  theme_pubr() +
  guides(alpha = "none", linetype = "none") +
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10), text = element_text(size = 10), legend.position = "none") +
  labs(x = expression('log'[10]*'T '*('After age correction'))) +
 scale_x_continuous(limits=c(0.24,0.48), n.breaks = 6) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2) 
# +
# annotate("text", x = xrng2[1], y = Inf, vjust = 1.1, label = lab4, size = 2)

```

```{r}
ggarrange(cutpoint_a_deaged, cutpoint_b_deaged, labels = c("A", "B"),  ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")
```

**FIGURE 2**

```{r}
fig_cutpoint_deaged_alt <- ggarrange(cutpoint_a, cutpoint_b,  cutpoint_a_deaged, cutpoint_b_deaged, labels = c("A", "C",  "B", "D"),  ncol = 2, nrow = 2, font.label = list(size = 11), common.legend = TRUE, legend = "bottom")

ggsave("fig_cutpoint_deaged_alt_adni.pdf", plot = fig_cutpoint_deaged_alt, width = 18, height = 12, units = "cm", device = "pdf")
# ggsave("fig_cutpoint_deaged_alt.png", plot = fig_cutpoint_deaged_alt, width = 9, height = 22, units = "cm", device = "png")
# ggsave("fig_cutpoint_deaged_alt.pdf", plot = fig_cutpoint_deaged_alt, dpi=1200, width = 9, height = 22, units = "cm", device = "pdf")

fig_cutpoint_deaged_alt
```

### Aging and pathological morphology alterations

Age groups:
```{r}
dados_hemi_v1_ADNI_agegroups <- dados_hemi_v1_ADNI %>%
  mutate(Age.group = ifelse(
    Age > 75,
    "76-86",
    ifelse((Age < 75 | Age == 75 & Age > 65 | Age == 65),
           "66-75",
           "")))


dados_hemi_v1_ADNI_agegroups %>%
  group_by(Age.group, Diagnostic) %>%
  summarise(N = n_distinct(SUBJ))

```

#### K
```{r}
summary(aov(K ~ Diagnostic:Age.group, data = dados_hemi_v1_ADNI_agegroups))
TukeyHSD(aov(K ~ Diagnostic:Age.group, data = dados_hemi_v1_ADNI_agegroups))

```

#### S
```{r}
summary(aov(S ~ Diagnostic:Age.group, data = dados_hemi_v1_ADNI_agegroups))
TukeyHSD(aov(S ~ Diagnostic:Age.group, data = dados_hemi_v1_ADNI_agegroups))
```

#### I
```{r}
summary(aov(I ~ Diagnostic:Age.group, data = dados_hemi_v1_ADNI_agegroups))
TukeyHSD(aov(I ~ Diagnostic:Age.group, data = dados_hemi_v1_ADNI_agegroups))

```

**Figure 3**

```{r}
mean_K_I_S <-
  dados_hemi_v1_ADNI %>% group_by(Diagnostic) %>% summarise(
    mean.K = mean(Knorm, na.rm = TRUE),
    SD_K = sd(Knorm, na.rm = TRUE),
    mean.I = mean(Inorm, na.rm = TRUE),
    SD_I = sd(Inorm, na.rm = TRUE),
    mean.S = mean(Snorm, na.rm = TRUE),
    SD_S = sd(Snorm, na.rm = TRUE),
    N_SUBJ = n_distinct(SUBJ)
    )
```

```{r K I S}
fig3a <- ggplot(mean_K_I_S, aes(x = mean.K, y = mean.S, color = Diagnostic)) +
  geom_point() +
  geom_line(group =1, color = "gray") +
  theme_pubr() +
  # theme(axis.title = element_text(size = 11),
  #   axis.text = element_text(size = 10), text = element_text(size = 10)) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)

fig3c <- ggplot(mean_K_I_S, aes(x = mean.K, y = mean.I, color = Diagnostic)) +
  geom_point() +
  geom_line(group =1, color = "gray") +
  theme_pubr() +
  # theme(axis.title = element_text(size = 11),
  #   axis.text = element_text(size = 10), text = element_text(size = 10)) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)
```

```{r}
mean_K_I_S_agegroups <-
  filter(dados_hemi_v1_ADNI_agegroups) %>% group_by(Diagnostic, Age.group) %>% summarise(
    mean.T = mean(logAvgThickness, na.rm = TRUE),
    SD_T = sd(logAvgThickness, na.rm = TRUE),
    mean.K = mean(Knorm, na.rm = TRUE),
    SD_K = sd(Knorm, na.rm = TRUE),
    mean.I = mean(Inorm, na.rm = TRUE),
    SD_I = sd(Inorm, na.rm = TRUE),
    mean.S = mean(Snorm, na.rm = TRUE),
    SD_S = sd(Snorm, na.rm = TRUE),
    N_SUBJ = n_distinct(SUBJ)
  )

```

```{r}
fig3b <- ggplot(mean_K_I_S_agegroups, aes(x = mean.K, y = mean.S, color = Diagnostic, shape = Age.group ))+
    geom_point() +
    geom_line(aes(group = Diagnostic)) +
    #geom_text(aes(label=Age.group), nudge_y = 0.1, size =3)+
    theme_pubr() + 
  # guides(color = "none") + 
    labs(shape = "Age", x = "K (normalized)", y = "S (normalized)") +
    # theme(axis.title = element_text(size = 11),
    #       axis.text = element_text(size = 10), text = element_text(size = 10)) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2) +
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin())

fig3d <- ggplot(mean_K_I_S_agegroups, aes(x = mean.K, y = mean.I, color = Diagnostic, shape = Age.group ))+
    geom_point() +
    geom_line(aes(group = Diagnostic)) +
    #geom_text(aes(label=Age.group), nudge_y = 0.1, size =3)+
    theme_pubr() +
  # guides(color = "none") + 
    labs(shape = "Age", x = "K (normalized)", y = "I (normalized)")  +
    # theme(axis.title = element_text(size = 11),
    #       axis.text = element_text(size = 10), text = element_text(size = 10)) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2) +
  theme(legend.position="bottom" , legend.box="vertical", legend.margin=margin())

fig3s <- ggarrange(fig3b, fig3d, labels = c("A", "B"), nrow=2, font.label = list(size = 11), common.legend = TRUE, legend = "bottom")

#fig3s <- ggarrange(fig3a, fig3b,fig3c, fig3d, labels = c("A", "B","C","D"), nrow=2, ncol=2, font.label = list(size = 11))

# ggsave("fig3s.pdf",fig3s, width = 9, height = 17.1, units = "cm", device = "pdf")
# ggsave("fig3s.png",fig3s, width = 9, height = 17.1, units = "cm", device = "png")

# ggsave("fig3s.pdf",fig3s, dpi = 1200, width = 18, height = 17.1, units = "cm", device = "pdf")
```

```{r figure3, fig.cap="\\label{fig:figure3}Morphological trajectory traced across the normalized independent components K, S,and I. We normalized the variable to the unity vectors providing comparable scale for the differences in both axes. (A, C) CTL as Control (blue, N~=~77) and the reference, MCI as Mild Cognitive Impairment (green, N~=~33) and AD for Alzheimer’s Disease (red, N~=~17). (B, D) Groups were divide in two subgroups, subjects with age between 65 and 75 years (circle, CTL N~=~67, MCI N~=~24, AD N~=~4) old and subjects with ages between 76 and 85 years old (triangle, CTL N~=~10, MCI N~=~9, AD N~=~9).", fig.height=6.73, fig.width=3.54 }
fig3s
```

##### Comparing IDOR vs ADNI

Age groups:
```{r}
dados_hemi_v1_agegroups <- dados_hemi_v1 %>%
  mutate(Age.group = ifelse(
    Age > 75,
    "76-86",
    ifelse((Age < 75 | Age == 75 & Age > 65 | Age == 65),
           "66-75",
           "")))

```

**Figure 3**

```{r}
mean_K_I_S <-
  dados_hemi_v1 %>% group_by(Sample, Diagnostic) %>% summarise(
    mean.K = mean(Knorm, na.rm = TRUE),
    SD_K = sd(Knorm, na.rm = TRUE),
    mean.I = mean(Inorm, na.rm = TRUE),
    SD_I = sd(Inorm, na.rm = TRUE),
    mean.S = mean(Snorm, na.rm = TRUE),
    SD_S = sd(Snorm, na.rm = TRUE),
    N_SUBJ = n_distinct(SUBJ)
    )
```

```{r}
fig3a <- ggplot(mean_K_I_S, aes(x = mean.K, y = mean.S, color = Diagnostic)) +
  geom_point() +
  geom_line(group =1, color = "gray") +
  theme_pubr() +
  # theme(axis.title = element_text(size = 11),
  #   axis.text = element_text(size = 10), text = element_text(size = 10)) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)

fig3c <- ggplot(mean_K_I_S, aes(x = mean.K, y = mean.I, color = Diagnostic)) +
  geom_point() +
  geom_line(group =1, color = "gray") +
  theme_pubr() +
  # theme(axis.title = element_text(size = 11),
  #   axis.text = element_text(size = 10), text = element_text(size = 10)) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)
```

```{r}
mean_K_I_S_agegroups <-
  filter(dados_hemi_v1_agegroups) %>% group_by(Sample, Diagnostic, Age.group) %>% summarise(
    mean.T = mean(logAvgThickness, na.rm = TRUE),
    SD_T = sd(logAvgThickness, na.rm = TRUE),
    mean.K = mean(Knorm, na.rm = TRUE),
    SD_K = sd(Knorm, na.rm = TRUE),
    mean.I = mean(Inorm, na.rm = TRUE),
    SD_I = sd(Inorm, na.rm = TRUE),
    mean.S = mean(Snorm, na.rm = TRUE),
    SD_S = sd(Snorm, na.rm = TRUE),
    N_SUBJ = n_distinct(SUBJ)
  )

```

```{r}
fig3b <- ggplot(mean_K_I_S_agegroups, aes(x = mean.K, y = mean.S, color = Diagnostic, shape = Age.group))+
    geom_point() +
    geom_line(aes(group = Diagnostic)) +
    #geom_text(aes(label=Age.group), nudge_y = 0.1, size =3)+
    theme_pubr() + 
  # guides(color = "none") + 
    labs(shape = "Age", x = "K (normalized)", y = "S (normalized)") +
    # theme(axis.title = element_text(size = 11),
    #       axis.text = element_text(size = 10), text = element_text(size = 10)) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2) +
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin()) +
  facet_grid(Sample ~ .)

fig3d <- ggplot(mean_K_I_S_agegroups, aes(x = mean.K, y = mean.I, color = Diagnostic, shape = Age.group ))+
    geom_point() +
    geom_line(aes(group = Diagnostic)) +
    #geom_text(aes(label=Age.group), nudge_y = 0.1, size =3)+
    theme_pubr() +
  # guides(color = "none") + 
    labs(shape = "Age", x = "K (normalized)", y = "I (normalized)")  +
    # theme(axis.title = element_text(size = 11),
    #       axis.text = element_text(size = 10), text = element_text(size = 10)) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2) +
  theme(legend.position="bottom" , legend.box="vertical", legend.margin=margin()) +
  facet_grid(Sample ~ .)

fig3s <- ggarrange(fig3b, fig3d, labels = c("A", "B"), nrow=2, font.label = list(size = 11), common.legend = TRUE, legend = "bottom")

#fig3s <- ggarrange(fig3a, fig3b,fig3c, fig3d, labels = c("A", "B","C","D"), nrow=2, ncol=2, font.label = list(size = 11))

ggsave("fig3s_adni.pdf",fig3s, width = 9, height = 17.1, units = "cm", device = "pdf")
# ggsave("fig3s.png",fig3s, width = 9, height = 17.1, units = "cm", device = "png")

# ggsave("fig3s.pdf",fig3s, dpi = 1200, width = 18, height = 17.1, units = "cm", device = "pdf")
```

```{r figure3alternativa, fig.height=6.73, fig.width=3.54 }
fig3s
```

## with age intervals

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dados_hemi_v1 %>%
  group_by(Diagnostic, Age_interval10) %>%
  summarise(
    N = n_distinct(SUBJ),
    age = paste(signif(mean(Age), 2), "±", signif(sd(Age), 2)),
    age_range = paste(signif(min(Age), 2), "; ", signif(max(Age), 2)),
    ESC = paste(signif(mean(ESC), 2), "±", signif(sd(ESC), 2))
  ) %>%
  kable(digits = 2) %>%
  kable_styling()
```


```{r}
mean_K_I_S_agegroups <-
  dados_hemi_v1 %>% group_by(Sample, Diagnostic, Age_interval10) %>% summarise(
    mean.T = mean(logAvgThickness, na.rm = TRUE),
    SD_T = sd(logAvgThickness, na.rm = TRUE),
    mean.K = mean(Knorm, na.rm = TRUE),
    SD_K = sd(Knorm, na.rm = TRUE),
    mean.I = mean(Inorm, na.rm = TRUE),
    SD_I = sd(Inorm, na.rm = TRUE),
    mean.S = mean(Snorm, na.rm = TRUE),
    SD_S = sd(Snorm, na.rm = TRUE),
    N_SUBJ = n_distinct(SUBJ)
  )

```

```{r}
fig3b_Alt <- ggplot(mean_K_I_S_agegroups, aes(x = mean.K, y = mean.S, color = Diagnostic, shape = Age_interval10))+
    geom_point() +
    geom_line(aes(group = Diagnostic)) +
  geom_errorbar(aes(ymin = mean.S - SD_S, ymax = mean.S + SD_S))  +
  geom_errorbarh(aes(xmin = mean.K - SD_K, xmax = mean.K + SD_K))  +
  geom_point(data = dados_hemi_v1_agegroups, aes(x = Knorm, y = Snorm, color = Diagnostic, shape = Age_interval10), alpha = 0.2) +
    #geom_text(aes(label=Age.group), nudge_y = 0.1, size =3)+
    theme_pubr() + 
  # guides(color = "none") + 
    labs(shape = "Age", x = "K (normalized)", y = "S (normalized)") +
    # theme(axis.title = element_text(size = 11),
    #       axis.text = element_text(size = 10), text = element_text(size = 10)) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2) +
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin()) +
  facet_grid(Sample ~ .)

fig3b_Alt
```

```{r}
fig3d_Alt <- ggplot(mean_K_I_S_agegroups, aes(x = mean.K, y = mean.I, color = Diagnostic, shape = Age_interval10))+
    geom_point() +
    geom_line(aes(group = Diagnostic)) +
    geom_errorbar(aes(ymin = mean.I - SD_I, ymax = mean.I + SD_I))  +
  geom_errorbarh(aes(xmin = mean.K - SD_K, xmax = mean.K + SD_K))  +
  geom_point(data = dados_hemi_v1_agegroups, aes(x = Knorm, y = Inorm, color = Diagnostic, shape = Age_interval10), alpha = 0.2) +
    #geom_text(aes(label=Age.group), nudge_y = 0.1, size =3)+
    theme_pubr() + 
  # guides(color = "none") + 
    labs(shape = "Age", x = "K (normalized)", y = "I (normalized)") +
    # theme(axis.title = element_text(size = 11),
    #       axis.text = element_text(size = 10), text = element_text(size = 10)) +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2) +
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin()) +
  facet_grid(Sample ~ .)
fig3d_Alt
```

```{r}
fig3s_Alt_allgroups <- ggarrange(fig3b_Alt, fig3d_Alt, labels = c("A", "B"),ncol = 2, nrow=1, font.label = list(size = 11), common.legend = TRUE, legend = "bottom")

fig3s_Alt_allgroups

ggsave("fig3s_Alt_allgroups_adni.pdf", fig3s_Alt_allgroups, width = 18, height = 18, units = "cm", device = "pdf")

```

```{r eval=FALSE, include=FALSE}

ggplot(dados_hemi_v1_ADNI, aes(x = Diagnostic, y = S, color = Diagnostic, fill = Diagnostic, alpha = 0.4)) +
  geom_boxplot() +
 stat_compare_means(method = "anova") +
  theme_pubr() +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)

aov_diag_S <- aov(S ~ Diagnostic, data = dados_hemi_v1_ADNI)
summary(aov_diag_S)
aov_diag_S_diag_TK <- TukeyHSD(aov_diag_S)
#aov_diag_S_diag_TK

aov_diag_S_diag_TK_t <- mutate(as.data.frame(aov_diag_S_diag_TK$Diagnostic), Contrast = c("MCI-AD", "CTL-AD", "CTL-MCI") , morphological_parameter = "S", Age_correction = "no", ROI = "hemisphere")

ggplot(dados_hemi_v1_ADNI, aes(x = Diagnostic, y = S_age_decay, color = Diagnostic, fill = Diagnostic, alpha = 0.4)) +
  geom_boxplot() +
 stat_compare_means(method = "anova") +
  theme_pubr() +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)

aov_diag_S_age_decay <- aov(S_age_decay ~ Diagnostic, data = dados_hemi_v1_ADNI)
summary(aov_diag_S_age_decay)
aov_diag_S_age_decay_diag_TK <- TukeyHSD(aov_diag_S_age_decay)
#aov_diag_S_age_decay_diag_TK

aov_diag_S_age_decay_diag_TK_t <- mutate(as.data.frame(aov_diag_S_age_decay_diag_TK$Diagnostic), Contrast = c("MCI-AD", "CTL-AD", "CTL-MCI") , morphological_parameter = "S", Age_correction = "yes", ROI = "hemisphere")

```
